Data-Driven Customization of Web Applications

### 2020-08-15 paper.tex
wrapping up visions

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Data-Driven Customization}

@@

We have developed a browser extension called Wildcard which uses web
scraping techniques to implement data-driven customization for existing
Web applications. In Section~\ref{sec:example}, we concretely
demonstrate the end user experience of using Wildcard to add features to
a news website.

In Section~\ref{sec:architecture}, we explain the system architecture of
data-driven customization. We focus on the \emph{table adapter}
abstraction, which allows many different types of underlying data to be
bidirectionally mapped to a table. We describe several types of table
adapters we've built in Wildcard, and also describe future adapters
supported by the general paradigm.

We have successfully used Wildcard to build customizations for 11
different websites that serve our own personal needs. In
Section~\ref{sec:reflections}, we present reflections from this process.
We outline the kinds of customizations we were able to build,
limitations we encountered, and some of the challenges of writing
scraping logic.

In Section~\ref{sec:visions}, we discuss some key themes from our work:

\begin{itemize}
\tightlist
\item
  \emph{Customization by direct manipulation}: We discuss further how
  data-driven customization relates to a gentle slope of customization.
  We suggest that an important point on this slope is the ability to
  customize an application by directly seeing and changing its data,
  rather than by writing imperative scripts.
\item
  \emph{Semantic wrappers}: Typically, tools that don't rely on official
  extension APIs resort to offering low-level APIs for customization.
  Instead, we propose a community-maintained library of semantic
  wrappers around existing applications, enabling end users to work with
  domain data rather than low-level representations.
\end{itemize}

Data-driven customization relates to existing work in many areas. Our
goals overlap with many software customization tools, and our methods
overlap with direct manipulation interfaces for working with structured
data, including visual database query systems and spreadsheets. We
explore these connections in Section~\ref{sec:related-work}.

@@

\hypertarget{decouple-data-from-applications}{%
\subsection{Decouple data from
applications}\label{decouple-data-from-applications}}

>>>>

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Data-Driven Customization of Web Applications}

@@

To explore this idea in a real context, we have developed a browser
extension called Wildcard that uses web scraping techniques to implement
data-driven customization for existing Web applications. We introduce
the tool with an example customization of Hacker News in
Section~\ref{sec:example}, and then describe the implementation in
{[}Section~\ref{sec:architecture}. In Section~\ref{sec:reflections}, we
present evidence that Wildcard can produce useful customizations, by
sharing reflections from customizing 11 different websites in ways that
met our own personal needs.

Wildcard is just an initial proof of concept of data-driven
customization. In Section~\ref{sec:visions}, we discuss our broader
vision for how this style of customization could change the relationship
between users and creators of software, focusing on three ideas:

\begin{itemize}
\tightlist
\item
  \emph{Decoupling data from applications}: On the modern Web, data is
  often stored in proprietary silos, limiting the agency of users to
  choose their applications and flexibly work with data. We propose
  data-driven customization as an incremental step towards a more
  decentralized architecture, where users gain more control over the
  storage, processing and display of information from web services.
\item
  \emph{Customization by direct manipulation}: We discuss how
  data-driven customization promotes a gentle slope of customization. We
  propose that an important point on this slope is the ability to
  customize an application by directly seeing and changing its data,
  rather than by writing imperative scripts.
\item
  \emph{Semantic wrappers}: Typically, tools that don't rely on official
  extension APIs resort to offering low-level APIs for customization.
  Instead, we propose a community-maintained library of semantic
  wrappers around existing applications, enabling end users to work with
  domain data rather than low-level representations.
\end{itemize}

In Section~\ref{sec:related-work} we discuss connections to related
work. Our goals overlap with software customization tools, and our
methods overlap with direct manipulation interfaces for working with
structured data, including visual database query systems and
spreadsheets.

@@

\hypertarget{decoupling-data-from-applications}{%
\subsection{Decoupling data from
applications}\label{decoupling-data-from-applications}}
